---
title: "Convex Project"
author: "Mykola Dereva"
date: "6/22/2022"
output: html_document
---

First of all lets load required libraries

```{r warning=FALSE}
rm(list = ls())

library(tidyverse)
library(here)
library(scales)
library(janitor)
library(ggrepel)

set.seed(42)
theme_set(theme_minimal())
```

# Load data

First, lets load all three datasets and clean them a bit so we can work with them.

## Cities data

Lets start with the set related to the biggest cities in the world

```{r}
cities <- read_csv(here(
  "data", "largest_cities.csv")) 


cities |> glimpse()
```

Definitely some transformation needed 

```{r}
cities <-  cities |> 
  # we don't need these columns
  select(-`Countries' Flags`,
         -`From 1950 - 2035`) |>  
  # convert to tidy format
  pivot_longer(-c(Country, City), 
               names_to = "Year",
               values_to = "Population") |> 
  # set appropriate data type 
  mutate(Year       = as.integer(Year),
         Population = as.integer(Population)) 

head(cities)
```

Data is loaded and tidy, lets try to do some summary statistic to better understand the data

```{r}
cities |> 
  group_by(Country, City) |> 
  summarise(na = sum(is.na(Population))) |> 
  arrange(desc(na))
```

For some Cities there are a lot of missing values we need to do something with them.

```{r}
cities |> 
  filter(City == "San Francisco")  
```

For example San Francisco. There is basically only one value for this city. Also, for very few cities there are data for > 2022 year, which I think is not reliable. So I will filter out those observations.

```{r}
cities <- cities |> 
  filter(Year <= 2020)
```

```{r}
cities |> 
  group_by(Country, City) |> 
  summarise(na = sum(is.na(Population))) |> 
  arrange(desc(na))
```

So far I don't know how I am going to use this data, so I will not do anything with those missing values.

## Biggest 10 Cities

Lets plot the population growth data for the 10 biggest cities for based for 2020 year

```{r}
top_10_cities <- cities |> 
  filter(Year == 2020) |> 
  arrange(desc(Population)) |> 
  head(10) |> 
  pull(City)
  
```

```{r}
cities |> 
  # select biggest cities
  filter(City %in% top_10_cities) |> 
  # set Cities as factor to plot colours properly
  mutate(City = factor(City, levels = top_10_cities)) |> 
  
  
  # plot
  ggplot(aes(x = Year, y = Population, 
             color = City)) + 
  geom_line(size = 0.75) +
  scale_y_continuous(labels = label_number_si()) +
  labs(title = "Dynamics of the Pupulation Growth of 10 Largest Cities")
```

The numbers are actually suspiciously high. I've checked on [wikipedia](https://en.wikipedia.org/wiki/Tokyo#Demographics) there was only 14 million of people living in Tokyo in 2020, but not > 30 million like on the graph. 

On the dataset documentation it is also not written about the nature of the data. 


## buildings data 

```{r}
bild <- read.csv(here("data", "tallest_buildings.csv")) |>
  clean_names()


bild |> glimpse()
```
I will have to do some data preparation 

```{r}
bild <- bild |> 
  # split height into two columns ft and m 
  separate(height, into = c("height_m", "height_ft"), sep = "\n/\n") |> 
  
  # extract number and convert to decimal 
  mutate(height_m =  str_extract(height_m,
                                 # some digits maybe dot and one digit
                                 regex("^[:digit:]*[:punct:]?[:digit:]")),
         height_m = as.double(height_m),
         
         height_ft = str_extract(height_ft, 
                                 # maybe digit maybe dot and some digits
                                 regex("^[:digit:]?[:punct:]?[:digit:]*")),
         height_ft = str_replace(height_ft, ",", "."),
         height_ft = as.double(height_ft), 
         
         # extract materials 
         material = str_extract(material, regex("[:lower:]*"))
         ) |> 
  rename(year = completion) |> 
  
  # drop one duplicated value
  filter(!duplicated(rank))


bild |> head()
```


### Relationship between year highest building 


```{r}
bild |> 
  arrange(year) |> 
  mutate(highest_building = cummax(height_m)) |> 
  
  # Plot
  ggplot(aes(x = year, y = height_m)) +
  geom_point(aes(color = material), size = 3, alpha = 0.5) +
  geom_line(aes(x = year, y = highest_building),
            alpha = 0.3, linetype = "dashed") +
  
  # add text 
  geom_text_repel(
    data = subset(bild, rank %in% c(1,10,17,23,49,87)),
      aes(label = name),
      size = 3,
      box.padding = 0.5,
      arrow = arrow(length = unit(0.010, "npc")),
      nudge_x = -2,
      nudge_y = 5,
      color = "grey50"
  ) +
  
  labs(title = "Evolution of the highest buildings",
       y = "Height in meters",
       x = "") 
  
  
```


## Load Countries Data

```{r}
count <- read.csv(here("data", "countries.csv"), dec = ",") |> 
  clean_names()

glimpse(count)
```




Lets explore some data a bit 

```{r}
count |> 
  summarise(across(everything(), ~ sum(is.na(.)))) |> 
  pivot_longer(cols = everything(),
               names_to = "column",
               values_to = "na_count")
```

We have some missing values


```{r}
count |> 
  count(region, sort = TRUE)
```


```{r}
count |> 
  mutate(region = as_factor(region) |> 
           fct_reorder(gdp_per_capita)) |> 

  ggplot(aes(x = gdp_per_capita, y = birthrate)) +
  geom_point(aes(col = region, size = population), alpha = 0.5) +

  scale_x_log10(label = dollar_format(accuracy = 1)) +
  geom_smooth(method = "lm", alpha = 0.3) + 
  scale_size(guide = FALSE) +
  
  labs(title = "Relationship between GDP Per Capita and Average Birthrate",
       x = "GDP Per Capita", 
       y = "Birthrate",
       caption = "Size of buble represent population size")
  
  


```




```{r}
sessionInfo()
```
